---
title: "HW4"
author: "Parijat"
date: "2/7/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("readr")
```
# Problem 1.  
```{r}
data(cheddar, package = "faraway")
fit <- lm(taste ~ . , data = cheddar)
dim(cheddar)
```

```{r}
anova(fit)
```

```{r}
?anova
```


1.1 Model tests whether the parameters of the model are significant.   

```{r}
lm1 = lm(taste ~ Acetic, data = cheddar)
lm2 = lm(taste ~ H2S, data = cheddar)
lm3 = lm(taste ~ Lactic, data = cheddar)
anova(lm1)
```

```{r}
anova(lm2)
```

```{r}
anova(lm3)
```
```{r}
?anova.lm
```

When we pass one model with all arguments, the tests are nested and the variance used for calculation is one for full model.


## Problem 2.  
```{r}
data(punting,package = "faraway")

dim(punting)
head(punting)
```

```{r}
lm_punting = lm(Distance ~ RStr+LStr+RFlex+LFlex,data = punting)
sm = summary(lm_punting)
sm
```
```{r}
#? summary(l)
qt(0.975,13-(4+1))
qt(0.025,13-(5))
sm$fstatistic
```

b) Since no values, are beyond the extreme t-values for 0.025 and 0.975 prob values for t-distribution, they are all significant
c) As p-value for over all f-statistic is 0.01 which is small, we can say that these predictors are important.

```{r}
mod_rl = lm(Distance ~ RStr+RFlex+LFlex,data = punting)
mod_ll = lm(Distance ~ LStr++RFlex+LFlex, data = punting)
anova(lm_punting,mod_rl)
anova(lm_punting,mod_ll)
```
p_value_right_leg =  0.726    
p_value_left_leg =  0.3228   

c) Since p-values are different for right leg and left leg,they may have different effect. Also, RSS comes to be different.

```{r}
anova(lm_punting,mod_rl,mod_ll)
```
```{r}
beta_rstr = lm_punting$coefficients[2]
best_lstr = lm_punting$coefficients[3]
```
 d) A joint confidence interval would be $$(X-K \hat\beta)^{T}(K(XX^T)^{-1}K^T)^{-1}(X-K\hat\beta)\leq m \hat\sigma^2 F_{p+1,n-p-1}( \alpha )$$
 where $$ K = \begin{bmatrix} 1 & 0 & 0 & 0 \\ 0 & 1 & 0 & 0  \end{bmatrix}$$
 The resulting equation, will give us an ellipsiod and it tells if the null is true, the parameters will lie inside the interval 95% times.
```{r}
mod_combined = lm(Distance ~ I(LStr+RStr)+ RFlex + LFlex,data=punting)
sm = summary(mod_combined)
#sm
```
 
 
 
```{r}
anova(lm_punting,mod_combined)  
```
 
 
 e) As F-statistic = 0.5804(very low) and p-value for reduced model is lesser than p-value of full-model, we may say that reduced model works atleast as good as full model.
 
 
```{r}
mod_rf = lm(Distance ~ RStr+RFlex+LStr,data = punting)
mod_lf = lm(Distance ~ LStr+RStr+LFlex, data = punting)
anova(lm_punting,mod_rf)
anova(lm_punting,mod_lf)
```

f) Since p-values are very different for the two models they may contribute differently

g)


```{r}
lm_hang = lm(Hang ~ RStr+LStr+RFlex+LFlex,data = punting)
summary(lm_hang)
```

h) No, as response variable is not same.

## 3. 
```{r}

#load("~/Desktop/Winter 2020/statistics/Statistics 552/HW3Data.RData",verbose=True)

#write.csv("~/Desktop/Winter 2020/statistics/Statistics 552/HW3Data.RData", "~/Desktop/hw3r.csv")
```

