---
title: "HW1"
author: "Parijat"
date: "1/16/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown
```{r libraries}

library("HistData")
library("Stat2Data")
library("ggplot2")
```



```{r Problem 1}

# 1.
emp = data.frame(
                id = c(1:4),
                Name = c("Alex","John","Wendy","JJ"),
                gender = factor(c("Male","Male","female","female") ),
              stringsAsFactors = FALSE
              )

# 2. 

emp_matrix = data.matrix(frame = emp)


# 3.
new_emp = cbind(emp)
new_emp[,3] = sapply(new_emp[,3], as.double)

# Factor categories given numeric values.

# 4.

data(GaltonFamilies, package = "HistData")
slr = lm(childHeight ~ midparentHeight,
data = GaltonFamilies)
cat("object slr is " , class(slr))
#slr is an object of type linear model

#slr$residuals

# 5. 

summary = summary(slr)

summary

summary$sigma


```

```{r Problem 2}

data(Fluorescence, package = "Stat2Data")

scatter.smooth( x= Fluorescence$Calcium, y = Fluorescence$ProteinProp)

slr_fl = lm(formula = ProteinProp ~ Calcium, data = Fluorescence)

#head(Fluorescence$Calcium)
#head(slr_fl$effects)
scatter.smooth(y = slr_fl$fitted.values, x= Fluorescence$Calcium )
```
The linear model may serve as a good fit for our objective. However, whether a model should be expected or not depends on how accurate our model is and what other good models we may have.


```{r}


num = Fluorescence$ProteinProp - 10^(-15)
logitp = log(num/(1-num))

inf = logitp[logitp == Inf]


scatter.smooth(y = logitp, x = Fluorescence$Calcium)

```
```{r}
slr_fl_log = lm(
          formula = logitp~calcium,
          data = data.frame(
            calcium = Fluorescence$Calcium,
            logitp = logitp
            ))


scatter.smooth(
  x= Fluorescence$Calcium,
  y= slr_fl_log$fitted.values
  )
```
```{r}
summary(slr_fl)
summary(slr_fl_log)

mean(slr_fl$residuals)
mean(slr_fl_log$residuals)
```
Second model fits better as it has larger P( > |t|)
```{r Problem 3}
data(GaltonFamilies, package = "HistData")
slr = lm(childHeight ~ midparentHeight,
data = GaltonFamilies)
summary = summary(slr)
std = summary$coefficients[2,2]
estimate = summary$coefficients[2,1]
n = length(slr$fitted.values)


summary
#estimate
#std

z_statistic = (estimate - 1)/std
z_statistic

prob = pnorm(z_statistic)

prob


```
## Since pvalue <= 0.025 for alpha = 5% null hypothesis test, we fail to reject the null hypothesis
## Y = 22.63624 + 0.63736*midparentHeight
## E[Y] = 22.6362 + 0.6376*E[midparentHeight]
## E[Y] = 22.6362 + 0.6376*72
```{r}
Y.bar = 22.6362 + 0.6376*72
Y.bar
```
# References
# 1. https://www.tutorialspoint.com/r


